---
title: "Final Project"
output: pdf_document
date: "2024-12-01"
---

## INTRODUCTION - This analysis explores the Ames, Iowa housing data using R's tidyverse package to understand the relationships between various features and the sale price of residential homes. It includes performing cleaning and transformations, generate summary statistics, create visualizations, and fit a linear regression model for prediction.


```{r}
### Reading the Data
# Load required libraries
library(tidyverse)
library(knitr)

file_path <- "/Users/surabhimetpally/Downloads/smetpal_house.csv"
housing_data <- read_csv(file_path)
head(housing_data)

```

```{r}
# Clean the data
cleaned_data <- housing_data %>%
  filter(RoofStyle != "Other", MoSold > 4) %>%
  mutate(SalePrice_scaled = SalePrice / 100000) %>%
  select(-YearRemodAdd, -HeatingQC)

# Display the first 10 rows and 6 columns
cleaned_data %>%
  select(1:6) %>%
  slice_head(n = 10) %>%
  kable()
```

```{r}
# Summary statistics
summary_stats <- cleaned_data %>%
  group_by(BsmtCond) %>%
  summarise(
    mean_SalePrice = mean(SalePrice, na.rm = TRUE),
    sd_SalePrice = sd(SalePrice, na.rm = TRUE),
    q1_SalePrice = quantile(SalePrice, 0.25, na.rm = TRUE),
    q3_SalePrice = quantile(SalePrice, 0.75, na.rm = TRUE),
    mean_GarageArea = mean(GarageArea, na.rm = TRUE),
    sd_GarageArea = sd(GarageArea, na.rm = TRUE),
    q1_GarageArea = quantile(GarageArea, 0.25, na.rm = TRUE),
    q3_GarageArea = quantile(GarageArea, 0.75, na.rm = TRUE),
    mean_GrLivArea = mean(GrLivArea, na.rm = TRUE),
    sd_GrLivArea = sd(GrLivArea, na.rm = TRUE),
    q1_GrLivArea = quantile(GrLivArea, 0.25, na.rm = TRUE),
    q3_GrLivArea = quantile(GrLivArea, 0.75, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the summary statistics
summary_stats %>%
  kable()

```

```{r}
# Scatter plot
ggplot(cleaned_data, aes(x = GarageArea, y = SalePrice, color = BsmtCond)) +
  geom_point() +
  theme_minimal() +
  labs(
    title = "SalePrice vs GarageArea",
    x = "Garage Area",
    y = "Sale Price",
    color = "Basement Condition"
  )
```

```{r}
# Fit the model
model <- lm(SalePrice ~ GarageArea + GrLivArea, data = cleaned_data)

# Model summary
summary(model)

```

```{r}
# Predict for new data
new_data <- tibble(
  GarageArea = c(500, 700),
  GrLivArea = c(1500, 2000)
)
predictions <- predict(model, newdata = new_data)
predictions
```

### Part 2

## For this section, I will explore and analyze an e-commerce dataset sourced from Kaggle. The dataset contains transaction details, including information about individual sales, products, and customer demographics. My goal is to use this data to answer two specific questions by performing both numerical and graphical analyses:
## What is the average quantity of items sold per invoice for each country?
## How does the total sales amount (calculated as Quantity * UnitPrice) vary across different countries?

```{r}
  
# Load required libraries
library(tidyverse)

# Dataset URL
url <- "https://www.kaggleusercontent.com/datasets/carrie1/ecommerce-data/download?datasetVersionNumber=1"

# Read the dataset into R
file_path <- "/Users/surabhimetpally/Downloads/e-commerce.data.csv"
data <- read_csv(file_path)
head(data)

```


```{r}
# Custom function to summarize numeric and categorical data
summarize_data <- function(data, columns = names(data), group_var = NULL) {
  # Select specified columns
  selected_data <- data %>% select(all_of(columns))
  
  # Initialize a list to store results
  results <- list()
  
  # Numeric columns: Mean, Variance, Quantiles
  numeric_cols <- selected_data %>% select(where(is.numeric))
  if (ncol(numeric_cols) > 0) {
    numeric_summary <- numeric_cols %>%
      summarise(across(everything(), list(
        mean = ~ mean(.x, na.rm = TRUE),
        variance = ~ var(.x, na.rm = TRUE),
        q05 = ~ quantile(.x, 0.05, na.rm = TRUE),
        q95 = ~ quantile(.x, 0.95, na.rm = TRUE)
      ), .names = "{.col}_{.fn}"))
    
    # Grouped numeric summaries if group_var is provided
    if (!is.null(group_var)) {
      numeric_summary <- numeric_cols %>%
        bind_cols(data %>% select(all_of(group_var))) %>%
        group_by(across(all_of(group_var))) %>%
        summarise(across(everything(), list(
          mean = ~ mean(.x, na.rm = TRUE),
          variance = ~ var(.x, na.rm = TRUE),
          q05 = ~ quantile(.x, 0.05, na.rm = TRUE),
          q95 = ~ quantile(.x, 0.95, na.rm = TRUE)
        ), .names = "{.col}_{.fn}"))
    }
    results$numeric_summary <- numeric_summary
  }
  
  # Categorical columns: Contingency tables
  categorical_cols <- selected_data %>% select(where(is.character))
  if (ncol(categorical_cols) > 0) {
    contingency_tables <- lapply(names(categorical_cols), function(col) {
      table(selected_data[[col]], useNA = "ifany")
    })
    names(contingency_tables) <- names(categorical_cols)
    results$contingency_tables <- contingency_tables
  }
  
  return(results)
}

```

```{r}
# Create a new column for total sales
data <- data %>%
  mutate(TotalSales = Quantity * UnitPrice)

# Apply the function: Summary statistics for TotalSales grouped by Country
summary_sales_by_country <- summarize_data(
  data,
  columns = c("TotalSales", "Quantity"),
  group_var = "Country"
)

# Apply the function: Summary for numeric columns only
summary_numeric <- summarize_data(data, columns = c("Quantity", "UnitPrice"))

# Apply the function: Contingency tables for categorical variables
summary_categorical <- summarize_data(data, columns = c("Country", "Description"))

# Display the results for grouped numeric data
summary_sales_by_country$numeric_summary

# Display contingency tables for categorical data
summary_categorical$contingency_tables

```

```{r}
## Question 1: Average Quantity of Items Sold per Invoice for the Top 5 Countries
# Plot average quantity sold per invoice for the first 5 countries
data %>%
  group_by(Country) %>%
  summarise(AverageQuantity = mean(Quantity, na.rm = TRUE)) %>%
  top_n(5, AverageQuantity) %>%  # Select top 5 countries by AverageQuantity
  ggplot(aes(x = reorder(Country, -AverageQuantity), y = AverageQuantity)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Average Quantity Sold per Invoice for the Top 5 Countries",
    x = "Country",
    y = "Average Quantity"
  )


```

```{r}
## Question 2: Total Sales by Country
# Plot total sales by country
data %>%
  group_by(Country) %>%
  summarise(TotalSales = sum(TotalSales, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(Country, -TotalSales), y = TotalSales)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(
    title = "Total Sales by Country",
    x = "Country",
    y = "Total Sales"
  )

```

